Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-01-11 src/ms.tex
paper edits

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{./figures/lamost-mcquillan.pdf}
    \includegraphics[width=\textwidth]{./figures/lamost-santos.pdf}
    \caption{Two-dimensional histograms indicating the number of Kepler target stars in the \teff–\prot plane for the \lamostmcq sample (top row) and the \lamostsan sample (bottom row). The effects of a simple cut in \logg to separate subgiants and dwarfs are shown in the middle and right columns. In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from its differentially rotating surface. The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature. The short-period pile-up is most clear in the \lamostmcq sample.}
    \label{fig:xmatch}
\end{figure*}

@@

It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation. However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well. As seen in Figure~\ref{fig:asteroseismic}, it is not clear that a pile-up exists for \teff~$\lesssim$~5800~K, particularly when examining the CKS and \citet{Hall2021} samples (see also \S\ref{subsec:extent}). However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K. We thus consider the explanation that WMB only operates for $\gtrsim$~1~\msun stars unlikely, and suggest that further work with a larger sample size could reveal the true \teff range of the pile-up more clearly.

We consider the most likely explanation for the observed gap between the Sun and the long-period pile-up to be systematic offsets in \teff. To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample. Notably, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift. Furthermore, as discussed in \S\ref{subsec:longperiod}, the \citet{Hall2021} sample is less subject to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here. As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period. We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases and small sample sizes.
}

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{./figures/lamost-mcquillan.pdf}
    \includegraphics[width=\textwidth]{./figures/lamost-santos.pdf}
    \caption{Two-dimensional histograms indicating the number of Kepler target stars in the \teff–\prot plane for the \lamostmcq sample (top row) and the \lamostsan sample (bottom row). The effects of a simple cut in \logg to separate subgiants and dwarfs are shown in the middle and right columns. In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from its differentially rotating surface. The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature. The short-period pile-up is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.}
    \label{fig:xmatch}
\end{figure*}

@@

Another possibility is that the long-period pile-up is only present for stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation. However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well. As shown in Figure~\ref{fig:xmatch}, it appears that the long-period pile-up does extend to temperatures as low as the Sun's, though this is not apparent from smaller samples such as the CKS sample or the \citet{Hall2021} asteroseismic sample (see also \S\ref{subsec:extent}). Kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, also suggests the pile-up exists at \teff as low as 5500~K or even 5000~K. We thus reject the explanation that WMB only operates for $\gtrsim$~1~\msun stars.

We consider the most likely explanation for the observed gap between the Sun and the long-period pile-up to be systematic offsets in \teff. To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample. Notably, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift. Furthermore, as discussed in \S\ref{subsec:longperiod}, the \citet{Hall2021} sample is less subject to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here. As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period. We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
}
